---
title: "Navigating the Fast Lane: A Case Study on Bike-Share's Speedy Success"
author: "Devi Nirfana H"
date: "`r Sys.Date()`"
output: "html_document"
---

This analysis is for case study 1 from the Google Data Analytics Certificate (Cyclistic). It's originally based on the case study "'Sophisticated, Clear, and Polished': Divvy and Data Visualization" written by Kevin Hartman(found here: <https://artscience.blog/home/divvy-dataviz-case-study>).

# Data Source and Limitations

[Download the Cyclistic trip data here.](https://divvy-tripdata.s3.amazonaws.com/index.html) The data has been made available by Motivate International Inc. under this [license](https://www.divvybikes.com/data-license-agreement).

The data does not include personally identifiable information (PII) about riders. This means that it is not possible to connect pass purchases to credit card numbers or to determine if casual riders live in the Cyclistic service area or if they have purchased multiple single passes. This limitation may impact the ability to make certain inferences about user behavior and preferences.

# Scenario

Cyclistic, a bike-share company in Chicago, aims to convert casual riders into annual members. Their marketing analyst team plans to analyze historical bike trip data to identify trends and develop targeted marketing campaigns to appeal to casual riders' needs.

As a junior data analyst, my focus is on unraveling the distinctive usage patterns between casual riders and annual members. The director of marketing emphasizes that the future success of Cyclistic hinges on maximizing annual memberships. In response to this directive, my team is diligently studying data to discern how these two user groups interact with Cyclistic bikes differently.

Our ultimate goal is to craft a targeted marketing strategy aimed at converting casual riders into committed annual members. However, before presenting our recommendations to Cyclistic executives, we understand the importance of substantiating our proposals with compelling data insights and professional visualizations.

I chose R for analyzing bike-share data due to its powerful statistical capabilities, extensive data manipulation tools, and rich ecosystem of data visualization packages.

# Import Data and Package Setup

```{r import packages}
library(tidyverse)
library(ggplot2)
library(lubridate)
```

```{r import data}
setwd("~/Documents/GitHub/R-project")

q1_2019 <- read_csv("./data/Divvy_Trips_2019_Q1.csv")
q2_2019 <- read_csv("./data/Divvy_Trips_2019_Q2.csv")
q3_2019 <- read_csv("./data/Divvy_Trips_2019_Q3.csv")
q4_2019 <- read_csv("./data/Divvy_Trips_2019_Q4.csv")
```

# Combining data into a single file

Before the data be combine, we have to identifying matching column names to facilitate data combination process.

```{r}
colnames(q1_2019)
```

```{r}
colnames(q2_2019)
```

```{r}
colnames(q3_2019)
```

```{r}
colnames(q4_2019)
```

Due to inconsistencies in column names and numbers, particularly in the Q2 2019 data, the column names will be standardized.

```{r renaming columns in q1_2019}
q1_2019 <- rename(q1_2019,
                  ride_id = trip_id,
                  bike_id = bikeid,
                  started_at = start_time,  
                  ended_at = end_time,
                  start_station_name = from_station_name, 
                  start_station_id = from_station_id,
                  end_station_name = to_station_name, 
                  end_station_id = to_station_id,
                  member_casual = usertype)
```

```{r renaming columns in q2_2019}
q2_2019 <- rename(q2_2019,
                   ride_id = "01 - Rental Details Rental ID",
                   bike_id = "01 - Rental Details Bike ID", 
                   started_at = "01 - Rental Details Local Start Time",  
                   ended_at = "01 - Rental Details Local End Time",  
                   start_station_name = "03 - Rental Start Station Name", 
                   start_station_id = "03 - Rental Start Station ID",
                   end_station_name = "02 - Rental End Station Name", 
                   end_station_id = "02 - Rental End Station ID",
                   member_casual = "User Type")
```

```{r renaming columns in q3_2019}
q3_2019 <- rename(q3_2019,
                   ride_id = trip_id,
                   bike_id = bikeid, 
                   started_at = start_time,  
                   ended_at = end_time,  
                   start_station_name = from_station_name, 
                   start_station_id = from_station_id, 
                   end_station_name = to_station_name, 
                   end_station_id = to_station_id, 
                   member_casual = usertype)
```

```{r renaming columns in q4_2019}
q4_2019 <- rename(q4_2019,
                   ride_id = trip_id,
                   bike_id = bikeid, 
                   started_at = start_time,  
                   ended_at = end_time,  
                   start_station_name = from_station_name, 
                   start_station_id = from_station_id, 
                   end_station_name = to_station_name, 
                   end_station_id = to_station_id, 
                   member_casual = usertype)
```

After Renaming the data then we review the datasets to uncover any discrepancies.

```{r}
str(q1_2019)
```

```{r}
str(q2_2019)
```

```{r}
str(q3_2019)
```

```{r}
str(q4_2019)
```

## changing data type

Apparently `ride_id` and `bike_id` was detected as numerical data type so we have to convert into character for ensure proper data stacking.

```{r converting columns in q1_2019}
q1_2019 <-  mutate(q1_2019, ride_id = as.character(ride_id),
                   bike_id = as.character(bike_id)) 
```

```{r converting columns in q2_2019}
q2_2019 <-  mutate(q2_2019, ride_id = as.character(ride_id),
                   bike_id = as.character(bike_id)) 
```

```{r converting columns in q3_2019}
q3_2019 <-  mutate(q3_2019, ride_id = as.character(ride_id),
                   bike_id = as.character(bike_id)) 
```

```{r converting columns in q4_2019}
q4_2019 <-  mutate(q4_2019, ride_id = as.character(ride_id),
                   bike_id = as.character(bike_id)) 
```

## Combining data into one big data frame.

```{r combine data}
all_trips <- bind_rows(q1_2019, q2_2019, q3_2019, q4_2019)
```

```{r}
colnames(all_trips)
```

The `birthyear` and `gender` will be removed from the dataset. The main focus was to find pattern and broader trends and insights related to bike sharing without compromising individual privacy.

The duration or `tripduration` column will recalculated later.

```{r removing unused columns}
all_trips <- all_trips %>%  
  select(c(ride_id, 
           started_at, 
           ended_at, 
           bike_id, 
           start_station_name, 
           start_station_id, 
           end_station_name, 
           end_station_id, 
           member_casual))
```

Since in this analysis we used member for subscriber and casual for a regular customer we will name the value inside the `member_casual` columns to be "Casual" for "Customer" and "Member" for "Subscriber".

```{r}
all_trips <-  all_trips %>% 
  mutate(member_casual = recode(member_casual
                                ,"Subscriber" = "Member"
                                ,"Customer" = "Casual"))
```

# Clean Up and Prepare data for analysis

Inspecting the new table that has been created.

```{r}
colnames(all_trips)
```

```{r}
dim(all_trips)
```

Upon inspection of the newly created table 'all_trips' we find that it comprises 10 columns and 3.818.004 rows.

```{r}
head(all_trips)  # See the first 6 rows of data frame. 
```

```{r}
tail(all_trips) # See the last 6 rows of data frame.
```

Examining the first and last six rows using the 'head' and 'tail' functions, respectively, reveals the structure and content of the data.

```{r}
str(all_trips)  #See list of columns and data types (numeric, character, etc)
```

```{r}
summary(all_trips)  #Statistical summary of data. Mainly for numerics
```

## Add more columns

Currently, the data is only available at the individual ride level, which limits the scope of aggregation and analysis. To expand the possibilities for data aggregation and gain a broader understanding of usage patterns, it is necessary to introduce additional columns containing information like day, month, and year. These additional columns will provide a temporal context for the ride data, enabling aggregation and analysis based on daily, monthly, or yearly trends.

```{r}
# Add columns that list the date, month, day, and year of each ride
# This will allow us to aggregate ride data for each month, day, or year  
# before completing these operations we could only aggregate at the ride level

all_trips$date <- as.Date(all_trips$started_at) #The default format is yyyy-mm-dd
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")

```

The `tripduration` columns will recalculated and renamed as `ride_length` representation of ride duration across all quarters.

```{r}
# the result will in minutes
all_trips$ride_length <- difftime(all_trips$ended_at,all_trips$started_at, units = "mins")
```

```{r}
str(all_trips)
```

Convert `ride_length` from Factor to numeric so we can run calculations on the data.

```{r}
is.factor(all_trips$ride_length)
```

```{r}
all_trips$ride_length <- as.numeric(as.character(all_trips$ride_length))
is.numeric(all_trips$ride_length)
```

## Remove Remove irrelevant data 

```{r}
summary(all_trips$ride_length)
```

The `tripduration` column contains negative values for some rides, including several hundred instances where the company removed bikes from circulation for quality control purposes. These rides will be removed from the dataset to ensure accurate analysis of ride duration.

To ensure data accuracy and consistency, a new version of the dataset (v2) is created by removing rows with negative ride lengths or rows where the bike ride originates from the HQ QR station.

```{r}
all_trips_clean <- all_trips[!(all_trips$start_station_name == "HQ QR" | all_trips$ride_length<0),]
```

# Time to analyse

## descriptive analysis

```{r}
summary(all_trips_clean$ride_length)
```

comparing casual and member user.

```{r}
aggregate(all_trips_clean$ride_length ~ all_trips_clean$member_casual, FUN = mean)
```

We can see that Casual member having longer average ride duration than Members.

```{r}
aggregate(all_trips_clean$ride_length ~ all_trips_clean$member_casual, FUN = median)
```

```{r}
aggregate(all_trips_clean$ride_length ~ all_trips_clean$member_casual, FUN = max)
```

```{r}
aggregate(all_trips_clean$ride_length ~ all_trips_clean$member_casual, FUN = min)
```

See the average ride time by each day for members vs casual users.

```{r}
aggregate(all_trips_clean$ride_length ~ all_trips_clean$member_casual + all_trips_clean$day_of_week, FUN = mean)
```

```{r}
# analyze ridership data by type and weekday
all_trips_clean %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>%
  group_by(member_casual, weekday) %>%
  summarise(number_of_rides = n(), average_duration = mean(ride_length)) %>%
  arrange(member_casual, weekday)	
```

## Visualize the data

```{r}
# Let's visualize the number of rides by rider type
viz1 <- all_trips_clean %>% mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(total_rides = n(),
            average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x = weekday, y = total_rides, fill = member_casual)) + 
  geom_text(
    aes(label = scales::comma(total_rides)),
    position = position_dodge(width = 0.9),  
    vjust = -0.5,  
    size = 1.8) +       
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(
    values = c("Casual" = "#f1a340", "Member" = "#998ec3"),
    breaks = c("Casual", "Member"),  # order of legend values
    labels = c("Casual Riders", "Annual Members")  # new legend value name
  ) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Casual Rider Vs. Annual Member",
    subtitle = "Ridership data comparison by member type on week days ",
    x = "Day of the Week",  # Change x-axis label
    y = "Total Rides",     # Change y-axis label
    fill = "Member Type")   # Change legend title
  
print(viz1)
```

```{r}
# Let's create a visualization for average duration
viz2 <- all_trips_clean %>% mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(total_rides = n(),
            average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  scale_fill_manual(
    values = c("Casual" = "#f1a340", "Member" = "#998ec3"),
    breaks = c("Casual", "Member"),  # order of legend values
    labels = c("Casual Riders", "Annual Members")  # new legend value name
  ) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Casual Rider Vs. Annual Member",
    subtitle = "Average ride duration based on member type",
    x = "Day of the Week",  # Change x-axis label
    y = "Average Duration (Min)",     # Change y-axis label
    fill = "Member Type")   # Change legend title

print(viz2)
```

# Answering business question

## How do annual members and casual riders use Cyclistic bikes differently?

```{r echo=FALSE}
print(viz1)
```

Casual Riders seem to prefer weekends, there total of 208,056 rides in Saturday and 170,173 rides in Sunday. While Annual Member tent to rides on weekdays.

```{r echo=FALSE}
print(viz2)
```

When it comes to average duration, Casual Riders have significantly longer ride duration than Annual Members, with daily rides averaging up to 60 minutes compared to less than 20 minutes for Annual Members.

## Why would casual riders buy Cyclistic annual memberships?

Based on the data Casual Riders are having longer duration than the Annual Member. Since then, it will be more cost-effective for them to purchase an annual membership.

Besides, after they purchases the annual membership they can use bike from any bike station without having to buy single riding ticket every time they want to use the bike.

This could be both cost-effective and time-saving for them in a long run.

## How can Cyclistic use digital media to influence casual riders to become members?

Cyclistic can collaborate with digital influencers who actively use bicycles for commuting or influencers promote sustainable transportation. By colaborating with these influential individuals, Cyclistic can reach a wider audience and raise awareness about their services. These influencers can also serve as potential leads for future marketing campaigns.

# Recommendation based on the analysis

1.  Cyclistic can work with digital influencer to reach a wider audience and raise awareness about their service. At the same time as the they reach a wider audience, Cyclistic can give casual riders a discount for their first membership or discount for the casual riders whose ride a certain number in a week.
2.  Partnering with local business to offer discount for Casual Riders whose rides bike from Cyclistic to get to the store or maybe co-working space. This would encourage them to consider the annual membership.
3.  Adding a point system or free beverages for the Members on a day in a week, this would increase riders to use the bike and persuade the casual rider to switch to annual member for the exclusive product or service.
